On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-12-02 manuscripts/PNAS/second review/resub/chazthm_pnas_resub.tex
explanations

\begin{proof}[Proof of Thm.~\ref{DeterministicUniquenessCorollary} for $k=1$]
Since the only 1-uniform hypergraph with the SIP is $[m]$, we have $\mathbf{x}_i = c_i \mathbf{e}_i$ for $c_i \in \mathbb{R} \setminus \{\mathbf{0}\}$, $i \in [m]$. In this case, we may use any $C_1 \geq 1/ \min_{\ell \in [m]} |c_{\ell}|$.

@@

\begin{definition}\label{FriedrichsDefinition}
For a collection of real subspaces $\mathcal{V} = \{V_i\}_{i=1}^\ell$, define $\xi^2 := 0$ when $|\mathcal{V}| = 1$, and otherwise:
\begin{align}\label{xi}
\xi^2(\mathcal{V}) := 1 -  \max \prod_{i=1}^{\ell-1} \sin^2  \theta \left(V_i, \cap_{j>i} V_j \right) ,
\end{align} 
%
where the maximum is taken over all orderings 
%\footnote{We modify the quantity in \cite{Deutsch12} in this way since the subspace ordering is irrelevant to our purpose.} 
of the $V_i$ and the angle $\theta \in (0,\frac{\pi}{2}]$ is defined implicitly as \cite[Def.~9.4]{Deutsch12}:
\begin{align*}
\cos{\theta(U,W)} := \max\left\{ |\langle \mathbf{u}, \mathbf{w} \rangle|: \substack{ \mathbf{u} \in U \cap (U \cap W)^\perp, \ \|\mathbf{u}\|_2 \leq 1 \\ \mathbf{w} \in W \cap (U \cap W)^\perp, \  \|\mathbf{w}\|_2 \leq 1 } \right\}.
\end{align*}
\end{definition}
Note that $\theta \in (0,\frac{\pi}{2}]$ implies $0 \leq \xi < 1$, and that $\xi(\mathcal{V}') \leq \xi(\mathcal{V})$ when $\mathcal{V}' \subseteq \mathcal{V}$.\footnote{We acknowledge the counter-intuitive property that $\theta =  \pi/2$ when $U \subseteq W$.}  
The constant $C_2$ in Lem.~\ref{MainLemma} is then:  
\begin{align}\label{Cdef2}
	C_2(\mathbf{A}, \mathcal{H}) := \frac{ (r+1) \max_{j \in [m]} \|\mathbf{A}_j\|_2}{ 1- \max_{\mathcal{G} \in {\mathcal{H} \choose r+1}} \xi( \bm{\mathcal{A}}_\mathcal{G} ) },
\end{align}
% C_2(\mathbf{A}, \mathcal{H}) := \frac{ (r+1) \max_{j \in [m]} \|\mathbf{A}_j\|_2}{ \min_{\mathcal{G} \in {\mathcal{H} \choose r} \cup {\mathcal{H} \choose r+1}} \xi( \{ \bm{\mathcal{A}}_S \}
which we remark yields a constant $C_1$ consistent with the case $k=1$ considered at the beginning of this section.\footnote{For $\mathbf{x}_i = c_i\mathbf{e}_i$, the denominator in \eqref{Cdef1} becomes $\min_{i \in [m]} |c_i| \|\mathbf{A}_i\|_2$, hence with $r=1$ we have $C_1 \geq 2 / \min_{i \in [m]} |c_i|$.}

The pragmatic reader should note that the explicit constants $C_1$ and $C_2$ are effectively computable: the quantity $L_k$ may be calculated as the smallest singular value of a certain matrix, while the quantity $\xi$ involves computing ``canonical angles'' between subspaces, which reduce again to an efficient singular value decomposition. There is no known fast computation of $L_k$ in general, however, since even $L_{k} > 0$ is NP-hard \cite{tillmann2014computational}; although fixing $k$ yields polynomial complexity. Moreover, calculating $C_2$ requires an exponential number of queries to $\xi$.  Thus, as presented, $C_1$ and $C_2$ are not efficiently computable.

@@

By Lem.~\ref{NonEmptyLemma}, the association $i \mapsto \cap \pi(\sigma(i))$ is an injective map $\overline \pi: J \to [\overline m]$ for some $J \subseteq [m]$ of size $\overline m - r(\overline m - m)$, and $\mathbf{B}_{\overline \pi(i)} \neq \mathbf{0}$ for all $i \in J$ since the columns of $\mathbf{B}_{\pi(S)}$ are linearly independent for all $S \in \mathcal{H}$. Letting $\overline \varepsilon := C_2 \varepsilon / \max_i \|\mathbf{A}_i\|_2$, it follows from \eqref{eqdim} and \eqref{last} that $d\left( \bm{\mathcal{A}}_i, \bm{\mathcal{B}}_{\overline \pi(i)} \right) \leq \overline \varepsilon$ for all $i \in J$. %Fixing $\overline \varepsilon = C_2\varepsilon$ and letting $c_i = \|\mathbf{A}_i\|_2^{-1}$, we thus have that for each $\mathbf{e}_i \in \mathbb{R}^m$ with $i \in J$ there exists some $\overline{c}_i \in \mathbb{R}$ such that $\|c_i\mathbf{A}\mathbf{e}_i - \overline{c}_i \mathbf{B}\mathbf{e}_{\overline \pi(i)}\|_2 \leq \overline \varepsilon < L_2(\mathbf{A}) \min_{i\in J} |c_i|$. But this is exactly the supposition in \eqref{1D}, and the result follows from the case $k=1$ in Sec.~\ref{DUT} applied to the submatrix $\mathbf{A}_J$. 
Setting $c_i := \|\mathbf{A}_i\|_2^{-1}$, by Def.~\ref{dDef} we have for all $i \in J$:
\begin{align*}
\min_{\overline c_i \in \mathbb{R}} \|c_i\bm{Ae}_i - \overline c_i \mathbf{Be}_{\overline \pi(i)} \|_2
\leq d\left( \bm{\mathcal{A}}_i, \bm{\mathcal{B}}_{\overline \pi(i)} \right)
\leq \overline \varepsilon
%&= \max_{\bm{u} \in \bm{\mathcal{A}}_i, \|\bm{u}\|_2 \leq 1} \text{dist}\left(\bm{u}, \bm{\mathcal{B}}_{\overline \pi(i)} \right) \\
%&\geq  \text{dist}\left(c_i\bm{Ae}_i, \mathbf{Be}_{\overline \pi(i)} \right) \\
%&= \min_{\overline c_i \in \mathbb{R}} \|c_i\bm{Ae}_i - \overline c_i \mathbf{Be}_{\overline \pi(i)} \|_2
\end{align*}
%
for $\overline \varepsilon < L_2(\mathbf{A})\min_{i \in [m]}|c_i|$. But this is exactly the supposition in \eqref{1D} (with $J$ in place of $[m]$), and the result thus follows from the case $k=1$ in Sec.~\ref{DUT} applied to the submatrix $\mathbf{A}_J$. 
\end{proof}

>>>>

\begin{proof}[Proof of Thm.~\ref{DeterministicUniquenessCorollary} for $k=1$]
Since the only 1-uniform hypergraph with the SIP is $[m]$, we have $\mathbf{x}_i = c_i \mathbf{e}_i$ for $c_i \in \mathbb{R} \setminus \{\mathbf{0}\}$, $i \in [m]$. In this case, we require that $C_1 \geq 1/ \min_{\ell \in [m]} |c_{\ell}|$.

@@

\begin{definition}\label{FriedrichsDefinition}
For a collection of real subspaces $\mathcal{V} = \{V_i\}_{i=1}^\ell$, define $\xi^2 := 0$ when $|\mathcal{V}| = 1$, and otherwise:
\begin{align}\label{xi}
\xi^2(\mathcal{V}) := 1 -  \max \prod_{i=1}^{\ell-1} \sin^2  \theta \left(V_i, \cap_{j>i} V_j \right) ,
\end{align} 
%
where the maximum is taken over all orderings 
%\footnote{We modify the quantity in \cite{Deutsch12} in this way since the subspace ordering is irrelevant to our purpose.} 
of the $V_i$ and the angle $\theta \in (0,\frac{\pi}{2}]$ is defined implicitly as \cite[Def.~9.4]{Deutsch12}:
\begin{align*}
\cos{\theta(U,W)} := \max\left\{ |\langle \mathbf{u}, \mathbf{w} \rangle|: \substack{ \mathbf{u} \in U \cap (U \cap W)^\perp, \ \|\mathbf{u}\|_2 \leq 1 \\ \mathbf{w} \in W \cap (U \cap W)^\perp, \  \|\mathbf{w}\|_2 \leq 1 } \right\}.
\end{align*}
\end{definition}
Note that $\theta \in (0,\frac{\pi}{2}]$ implies $0 \leq \xi < 1$, and that $\xi(\mathcal{V}') \leq \xi(\mathcal{V})$ when $\mathcal{V}' \subseteq \mathcal{V}$.\footnote{We acknowledge the counter-intuitive property that $\theta =  \pi/2$ when $U \subseteq W$.}  
The constant $C_2$ in Lem.~\ref{MainLemma} is then:  
\begin{align}\label{Cdef2}
	C_2(\mathbf{A}, \mathcal{H}) := \frac{ (r+1) \max_{j \in [m]} \|\mathbf{A}_j\|_2}{ 1- \max_{\mathcal{G} \in {\mathcal{H} \choose r+1}} \xi( \bm{\mathcal{A}}_\mathcal{G} ) },
\end{align}
% C_2(\mathbf{A}, \mathcal{H}) := \frac{ (r+1) \max_{j \in [m]} \|\mathbf{A}_j\|_2}{ \min_{\mathcal{G} \in {\mathcal{H} \choose r} \cup {\mathcal{H} \choose r+1}} \xi( \{ \bm{\mathcal{A}}_S \}
which we remark yields a constant $C_1$ consistent with what is required for the case $k=1$ considered at the beginning of this section.\footnote{For $\mathbf{x}_i = c_i\mathbf{e}_i$, the denominator in \eqref{Cdef1} becomes $\min_{i \in [m]} |c_i| \|\mathbf{A}_i\|_2$, hence with $r=1$ we have $C_1 \geq 2 / \min_{i \in [m]} |c_i|$.}

The pragmatic reader should note that the explicit constants $C_1$ and $C_2$ are effectively computable: the quantity $L_k$ may be calculated as the smallest singular value of a certain matrix, while the quantity $\xi$ involves computing ``canonical angles'' between subspaces, which reduce again to an efficient singular value decomposition. There is no known fast computation of $L_k$ in general, however, since even $L_{k} > 0$ is NP-hard \cite{tillmann2014computational}; although fixing $k$ yields polynomial complexity. Moreover, calculating $C_2$ requires an exponential number of queries to $\xi$ unless $r$ is held fixed, too (e.g., the `cyclic order' hypergraphs described above have $r=k$).  Thus, as presented, $C_1$ and $C_2$ are not efficiently computable.

@@

By Lem.~\ref{NonEmptyLemma}, the association $i \mapsto \cap \pi(\sigma(i))$ is an injective map $\overline \pi: J \to [\overline m]$ for some $J \subseteq [m]$ of size $\overline m - r(\overline m - m)$, and $\mathbf{B}_{\overline \pi(i)} \neq \mathbf{0}$ for all $i \in J$ since the columns of $\mathbf{B}_{\pi(S)}$ are linearly independent for all $S \in \mathcal{H}$. Letting $\overline \varepsilon := C_2 \varepsilon / \max_i \|\mathbf{A}_i\|_2$, it follows from \eqref{eqdim} and \eqref{last} that $d\left( \bm{\mathcal{A}}_i, \bm{\mathcal{B}}_{\overline \pi(i)} \right) \leq \overline \varepsilon$ for all $i \in J$. %Fixing $\overline \varepsilon = C_2\varepsilon$ and letting $c_i = \|\mathbf{A}_i\|_2^{-1}$, we thus have that for each $\mathbf{e}_i \in \mathbb{R}^m$ with $i \in J$ there exists some $\overline{c}_i \in \mathbb{R}$ such that $\|c_i\mathbf{A}\mathbf{e}_i - \overline{c}_i \mathbf{B}\mathbf{e}_{\overline \pi(i)}\|_2 \leq \overline \varepsilon < L_2(\mathbf{A}) \min_{i\in J} |c_i|$. But this is exactly the supposition in \eqref{1D}, and the result follows from the case $k=1$ in Sec.~\ref{DUT} applied to the submatrix $\mathbf{A}_J$. 
Setting $c_i := \|\mathbf{A}_i\|_2^{-1}$ so that $\|c_i\mathbf{Ae}_i\|_2 = 1$, by Def.~\ref{dDef} we have for all $i \in J$:
\begin{align*}
\min_{\overline c_i \in \mathbb{R}} \|c_i\mathbf{Ae}_i - \overline c_i \mathbf{Be}_{\overline \pi(i)} \|_2
\leq d\left( \bm{\mathcal{A}}_i, \bm{\mathcal{B}}_{\overline \pi(i)} \right)
\leq \overline \varepsilon
%&= \max_{\bm{u} \in \bm{\mathcal{A}}_i, \|\bm{u}\|_2 \leq 1} \text{dist}\left(\bm{u}, \bm{\mathcal{B}}_{\overline \pi(i)} \right) \\
%&\geq  \text{dist}\left(c_i\bm{Ae}_i, \mathbf{Be}_{\overline \pi(i)} \right) \\
%&= \min_{\overline c_i \in \mathbb{R}} \|c_i\bm{Ae}_i - \overline c_i \mathbf{Be}_{\overline \pi(i)} \|_2
\end{align*}
%
for $\overline \varepsilon < L_2(\mathbf{A})\min_{i \in [m]}|c_i|$. But this is exactly the supposition in \eqref{1D}, with $J$ and $\overline \varepsilon$ in place of $[m]$ and $\varepsilon$, respectively. Thus, applying the arguments of the case $k=1$ in Sec.~\ref{DUT} to the submatrix $\mathbf{A}_J$, we have $\|\mathbf{A}_j - \mathbf{B}_{\overline J}\mathbf{PD}_j\|_2 \leq \overline  \varepsilon / |c_j|  \leq C_2 \varepsilon$ for $j \in J$.
\end{proof}
